### 2021-07-03 rad-icfp-2021/relwork.tex
bit more related work

@@

\noindent \textbf{Limitations and future work}
\begin{itemize}
    \item memory access for cost function 
    \item parallelism
    \item dedicated tensors for ML, and BLaS considerations
    \item implementation and benchmark
    \item conditionals
    \item recursion
    \item general HO function
    \item more 2nd order functions, though can be derived by hand once and then it's ok
    \item more dynamic arrays
    \item more dynamic control flow
    \item all store vs all recompute reverse-mode, and advanced checkpointing
    \item numerical accuracy
    \item fancier types and datatypes (eg dependent, adt) 
    \item iterated ad and higher derivatives
    \item use in Probabilistic programming
    \item non smooth and partial functions, although by simply following the ref book it's as usual in AD
    \item sparsity and data manipulation
    \item other optimisations like implicit function theorem
\end{itemize}

\MH{useful links (Swift is close to our project, and the doc is very readable):}
\begin{verbatim}https://docs.google.com/document/d/1_BirmTqdotglwNTOcYAW-ib6mx_jl-gH9Dbg4WmHZh0/edit# \end{verbatim}
\begin{verbatim}
	https://github.com/apple/swift/blob/master/docs/DifferentiableProgramming.md#approaches-to-automatic-differentiation
\end{verbatim}

\MH{understand related work sentence from Diff. Curry paper: The idea of using closures as back-propagators is receiving recent attention. For example Julia Zy- gote [15] and Swift AD adopts this design. Other recent work follows similar ideas [27, 26] but is using meta-programming as an implementation technique.}

>>>>

Good recent surveys: \cite{van2018automatic,baydin2017automatic}

@@

\noindent \textbf{Cost function and memory accesses}
Our cost model \ref{sub:costModel} is simple but has limits. 
In particular, it does not capture well cost in term of
parallel computing, and memory management (cache optimisation is one of best way to optimise such linear algebra intensive computations).
For instance, in the earlier days of automatic differentiation, 
advanced checkpointing techniques were used to avoid storing all the intermediate values just to save memory 
and try avoiding saturation of RAM. 
These problems are harder to analyse and eventually it seems to come down to a lot of experimenting and benchmarking.

\noindent \textbf{Parallelism and GPU}
We tried to locally preserve parallelism as much as possible, 
and having a pure transformation should help with found parallelism. 
It would be interesting to implement on transformations and see how they perform on GPU.
It is well-kwown that the addition of references provoked by the standard reverse-mode transformation
restricts a lot of optimisations and it is not hard to find examples in C where the compiler does not detect simple optimisations 
because of the references introduced by reverse-mode.

\noindent \textbf{implementation, BLaS, benchmark}

\noindent \textbf{Higher-order functions}
Matthijs'work, use of defunctionalization, question efficiency and usefulness, 
interested in functions like root of function or minimal of a function. Implicit function theorem

\noindent \textbf{More array primitives}
filter, flatten, gather, scanl, ...
can be derived by hand once and then it's ok

\noindent \textbf{Recursion}

\noindent \textbf{Higher derivatives}

\noindent \textbf{Sparse data and matrix representation}

\begin{itemize}
    \item fancier types and datatypes (eg dependent, adt) 
    \item use in Probabilistic programming
    \item other optimisations like implicit function theorem
\end{itemize}

% \MH{useful links (Swift is close to our project, and the doc is very readable):}
% \begin{verbatim}https://docs.google.com/document/d/1_BirmTqdotglwNTOcYAW-ib6mx_jl-gH9Dbg4WmHZh0/edit# \end{verbatim}
% \begin{verbatim}
% 	https://github.com/apple/swift/blob/master/docs/DifferentiableProgramming.md#approaches-to-automatic-differentiation
% \end{verbatim}

\MH{understand related work sentence from Diff. Curry paper: The idea of using closures as back-propagators is receiving recent attention. 
    For example Julia Zygote \cite{innes2019zygote} and Swift AD \cite{wei2018first} chose this design. 
    Other recent work follows similar ideas [27, 26] but is using meta-programming as an implementation technique.}
